---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Model Comparison {#model-comparison-int}

## Formalize Models

We consider Zero Inflated Negative Binomial Mixed-Effects models with only the role of gender as a fixed effect and children's classroom ID as a random effect for $p$. Whereas, considering $\mu$, we define the same models as in the analysis of externalizing problems. Using R formula syntax, we have
```{r, echo = TRUE, eval=FALSE}
# formula for p (same for all models)
p ~ gender + (1|ID_class)

# formula for mu

# fit_int_zero
mu ~ gender + (1|ID_class)

# fit_int_mother
mu ~ gender + mother + (1|ID_class)

# fit_int_additive
mu ~ gender + mother + father + (1|ID_class)

# fit_int_inter
mu ~ gender + mother * father + (1|ID_class)
```

## AIC and BIC Results

AIC and BIC values together with their relative weights are computed are reported in Table~\@ref(tab:table-AIC-BIC-weights-int).

```{r table-AIC-BIC-weights-int, cache=TRUE}
get_table_AIC_BIC(AIC_weights = AIC_weights_int,
                  BIC_weights = BIC_weights_int,
                  problem = "int",
                  format = params$format,
                  path_img = "images/")
```

According to AIC, the most likely model is `fit_int_mother` (`r perc_model_int("fit_int_mother", "AIC")`\%) and the second most likely model is `fit_int_additive` (`r perc_model_int("fit_int_additive", "AIC")`\%) given the data and the set of models considered. According to BIC, instead, `fit_int_zero` and `fit_int_mother` models have almost the same probability, `r perc_model_int("fit_int_zero", "BIC")`\% and `r perc_model_int("fit_int_mother", "BIC")`\% respectively.

We can say that there is evidence in favour of the role of mother attachment but probably this effect is small.

## Selected Model

Results of analysis of deviance for  model `fit_int_mother` are reported below.
```{r, echo =TRUE, cache=TRUE}
car::Anova(fit_int_mother)
```
Results confirm a statistically significant effect of mother attachment. The model summary is reported below.
```{r, echo =TRUE, cache=TRUE}
summary(fit_int_mother)
```
To evaluate the effect of mother attachment, the marginal predicted values are presented in Figure~\@ref(fig:plot-comparison-effects-int). Not that the marginal predicted values for gender are averaged over mother attachment. Whereas, the marginal predicted values for mother attachment are averaged over gender.
```{r plot-comparison-effects-int, cache=TRUE, fig.asp=.65, message=FALSE, fig.cap="Marginal predicted values according to gender and mother attachment ($n_{subj} = 847$)."}
get_plot_zinb(model = fit_int_mother, attachment = "mother", gender = FALSE)
```

Post-hoc tests are run, considering pairwise comparisons and adjusting *p*-values according to multivariate *t*-distribution. Results are reported below,
```{r, cache=TRUE, echo = TRUE}
emmeans::contrast(emmeans::emmeans(fit_int_mother, specs = ~ mother ),
                  "pairwise", adjust = "mvt")
```

Results indicate that Fearful and Anxious children have more problems than Secure children. Moreover, also the difference between Avoidant and Fearful children is significant.

To evaluate the fit of the model to the data, we computed the *Marginal* $R^2$ and the *Conditional* $R^2$.

```{r, echo=TRUE, message=FALSE, cache=TRUE}
performance::r2(fit_int_mother)
```

We can see that the actual variance explained by fixed effects is 3\%, not that much.

#### Conclusions {-}

Considering attachment theoretical perspectives, results indicate only the role of mother attachment so we can support the **Monotropy Theory**. Note, however, that the compared models contain no information regarding the expected direction of the effects but we only include/exclude predictors.

